---
title: "Matched sample Creation and Diagnostics"
author: "Joel Earlywine"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

---
title: "Matched Sample Creation and Diagnostics"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(MatchIt)
library(cobalt)
library(ggplot2)

source("01_packages.R")
source("02_load_data.R")

df1_original <- df1
```

# Matching specification

```{r}
match_model <- matchit(
  treated ~ x01 + x02 + x03 + x04 + x05 + x06 + x07 + x09,
  data = df1_original,
  method = "nearest",
  distance = "glm",
  ratio = 1
)

match_model
```

# Matching summary

```{r}
summary(match_model)
```

# Balance diagnostics table

```{r}
balance_table <- cobalt::bal.tab(match_model)

balance_table
```

# Love plot

```{r, fig.width=8, fig.height=6}
cobalt::love.plot(
  match_model,
  binary = "std",
  thresholds = c(m = 0.1),
  abs = TRUE
)
```

# Propensity score overlap before matching

```{r, fig.width=8, fig.height=5}
plot(
  match_model,
  type = "density",
  interactive = FALSE
)
```

# Propensity score jitter plot

```{r, fig.width=8, fig.height=5}
plot(
  match_model,
  type = "jitter",
  interactive = FALSE
)
```

# Extract matched sample

```{r}
df1_matched <- match.data(match_model)

df1_matched_clean <- df1_matched %>%
  mutate(
    matched_sample = TRUE,
    matching_method = "nearest_neighbor_1to1",
    propensity_score = distance,
    matched_wgt = weights,
    matched_set_id = subclass
  ) %>%
  select(
    id,
    matched_sample,
    matched_set_id,
    matching_method,
    propensity_score,
    matched_wgt,
    everything(),
    -distance,
    -weights,
    -subclass
  )

dim(df1_matched_clean)
```

# Variable names in matched dataset

```{r}
names(df1_matched_clean)
```

# Matched sample treatment counts

```{r}
df1_matched_clean %>%
  count(treated)
```

# Original vs matched sample size

```{r}
sample_size_comparison <- data.frame(
  sample = c("Original", "Matched"),
  rows = c(nrow(df1_original), nrow(df1_matched_clean)),
  treated_n = c(
    sum(df1_original$treated == TRUE),
    sum(df1_matched_clean$treated == TRUE)
  ),
  control_n = c(
    sum(df1_original$treated == FALSE),
    sum(df1_matched_clean$treated == FALSE)
  )
)

sample_size_comparison
```

# Outcome comparison in matched sample

```{r}
matched_outcome_summary <- df1_matched_clean %>%
  group_by(treated) %>%
  summarise(
    n = n(),
    mean_birthweight = mean(y01, na.rm = TRUE),
    sd_birthweight = sd(y01, na.rm = TRUE),
    min_birthweight = min(y01, na.rm = TRUE),
    max_birthweight = max(y01, na.rm = TRUE),
    .groups = "drop"
  )

matched_outcome_summary
```

# Weighted matched-sample regression

```{r}
matched_lm <- lm(
  y01 ~ treated,
  data = df1_matched_clean,
  weights = matched_wgt
)

summary(matched_lm)
```

# Save matched sample

```{r}
saveRDS(
  df1_matched_clean,
  file = "cattaneo2_matched_sample.rds"
)

write.csv(
  df1_matched_clean,
  file = "cattaneo2_matched_sample.csv",
  row.names = FALSE
)

cat("Saved files:\n")
cat("- cattaneo2_matched_sample.rds\n")
cat("- cattaneo2_matched_sample.csv\n")
```

# Final summary

```{r}
# Comprehensive variable diagnostics
summarytools::dfSummary(
  df1_matched_clean,
  graph.col = FALSE
)
```

# Session information

```{r}
sessionInfo()
```